What Boolean Algebras Can Teach Us About Weighted Model Counting

### 2020-04-13 paper.tex
Declutter + update/improve

\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\newtheorem{proposition}{Proposition}
\newtheorem{corollary}{Corollary}
\newtheorem{conjecture}{Conjecture}
\theoremstyle{definition}
\newtheorem{definition}{Definition}
\newtheorem{example}{Example}
\theoremstyle{remark}
\newtheorem*{remark}{Remark}

\definecolor{color1}{HTML}{fbb4ae}
\definecolor{color2}{HTML}{b3cde3}
\definecolor{color3}{HTML}{ccebc5}
\definecolor{color4}{HTML}{decbe4}

@@

% Feedback: Can you say something here about factorized vs non-factorized
% weight function definitions? That is, factorized is when w maps literals to
% R_>=0, non-factorized is when w maps models to R_>=0 and show:
% a) come up with nice example when non-factorized weights are intuitive
% b) what if weight functions are negative/complex?
% c) clarify that the factorized definition have is w.r.t. models, in case some
% one gets confused [It doesn't have to be, if the BA is not free -- P.]

% Feedback 2: WMC-Computable -> WMC-Definable

@@

% Feedback 2: I did a high level read to get the big picture but found
% this discussion on ideals distracting. Specifically, w.r.t. the main claims
% when do these things get used?

% TODO:
% 1) theorem: a pushout of free algebras is a free algebra. Same for
% coproducts. This extends to coproducts because they are a special case of
% pushouts.
% TODO for later:
% 2) curly brackets or parentheses for families? Maybe curly brackets
% 3) make up my mind about a,b vs x,y and stick to it (maybe x,y?)
% 4) add parentheses around quotients (e.g., (a+b)/I).
% 5) perhaps reorder the section of preliminaries into paragraphs, i.e., a
% paragraph for order, for homomorphisms, for ideals and quotients. This would
% take up less space.

@@

% TODO: Feedback 2: Explain with examples: models = elements [atoms] of algebra

@@

For clarity and succinctness, we will occasionally use three other operations
that can be defined using the original three\footnote{We use $+$ to denote
  symmetric difference because it is the additive operation of a Boolean ring.}:
\begin{align*}
  a \to b &= \neg a \lor b, \\
  a \leftrightarrow b &= (a \land b) \lor (\neg a \land \neg b), \\
  a + b &= (a \land \neg b) \lor (\neg a \land b).
\end{align*}
We can also define a partial order $\le$ on $\mathbf{B}$ as $a \le b$ if $a = b
\land a$ (or, equivalently, $a \lor b = b$) for $a, b \in \mathbf{B}$.
Furthermore, let $a < b$ denote $a \le b$ and $a \ne b$. For the rest of this
paper, let $\mathbf{B}$ refer to the BA $(\mathbf{B}, \land, \lor, \neg, 0, 1)$.
For any $S \subseteq \mathbf{B}$, we write $\bigvee S$ for $\bigvee_{x \in S} x$
and call it the \emph{supremum} of $S$. Similarly, $\bigwedge S = \bigwedge_{x
  \in S} x$ is the \emph{infimum}. By convention, $\bigwedge \emptyset = 1$ and
$\bigvee \emptyset = 0$. For any $a, b \in \mathbf{B}$, we say that $a$ and $b$
are \emph{disjoint} if $a \land b = 0$.

@@

%\begin{definition}[\cite{givant2008introduction}] \label{def:filter}
%  A \emph{(Boolean) filter} is a non-empty subset $F \subseteq \mathbf{B}$ such
%  that
%  \begin{itemize}
%  \item $f \land g \in F$ for all $f, g \in F$,
%  \item $f \lor a \in F$ for all $f \in F$ and $a \in \mathbf{B}$.
%  \end{itemize}
%  For any $p \in \mathbf{B}$, the \emph{principal filter of $p$} is the
%  intersection $I$ of all filters that contain $p$. It also has the property
%  that $p \le a$ for all $a \in I$.
%\end{definition}
%\begin{definition}[\cite{sikorski1969boolean}] \label{def:quotient}
%  Let $F \subseteq \mathbf{B}$ be a filter. The \emph{quotient} $\mathbf{B}/F$
%  is a BA of equivalence classes of elements of $\mathbf{B}$ with respect to the
%  equivalence relation
%  \[
%    a \sim b \qquad \iff \qquad a \to b \in F \quad \text{and} \quad b \to a \in F
%  \]
%  for all $a, b \in \mathbf{B}$. Elements of $\mathbf{B}/F$ are usually denoted
%  by $a/F$ (for some $a \in \mathbf{B}$) with the understanding that if $b \sim
%  a$ (for some $b \in \mathbf{B}$), then $b/F = a/F$.
%\end{definition}

\begin{definition}[\cite{givant2008introduction}] \label{def:ideal}
  An \emph{ideal} is a non-empty subset $I \subseteq \mathbf{B}$ such that
  \begin{itemize}
  \item $i \lor j \in I$ for all $i, j \in I$;
  \item $i \land a \in I$ for all $i \in I$ and $a \in \mathbf{B}$.
  \end{itemize}
  For any $p \in \mathbf{B}$, the \emph{principal ideal of $p$}---denoted by
  $(p)$---is the smallest ideal that contains $p$. It can also be expressed as
  $(p) = \{ a \in \mathbf{B} \mid a \le p \}$.
\end{definition}

\begin{definition}[\cite{givant2008introduction}] \label{def:quotient}
  Let $I$ be an ideal in $\mathbf{B}$. The \emph{quotient algebra of
    $\mathbf{B}$ modulo the ideal $I$} $\mathbf{B}/I$ is a BA
  of equivalence classes of elements of $\mathbf{B}$ with respect to the
  equivalence relation
  \[
    a \sim b \quad \iff \quad a + b \in I
  \]
  for all $a, b \in \mathbf{B}$. Elements of $\mathbf{B}/I$ are usually denoted
  by $a/I$ (for some $a \in \mathbf{B}$) with the understanding that if $b \sim
  a$ (for some $b \in \mathbf{B}$), then $b/I = a/I$. The three algebraic
  operations on $\mathbf{B}/I$ are defined as
  \begin{align*}
    a/I \land b/I &= a \land b / I, \\
    a/I \lor b/I &= a \lor b / I, \\
    \neg(a/I) &= (\neg a)/I.
  \end{align*}
\end{definition}

@@

\begin{lemma}[\cite{givant2008introduction}] \label{lemma:canonical_homomorphism}
  Let $I \subseteq \mathbf{B}$ be an ideal. The map $f\colon \mathbf{B} \to
  \mathbf{B}/I$ defined by $f(x) = x/I$ is a homomorphism.
\end{lemma}

@@

\section{WMC as a Measure}

@@

\begin{figure}
  \[
    \begin{tikzcd}
      & & & & & \colorbox{color3}{$1$} \ar[dlll,dash,gray] \ar[dl,dash,gray]
      \ar[dr,dash,gray] \ar[drrr,dash,gray] & & & \\
      & & \colorbox{color3}{$p \lor q$} & & \colorbox{color3}{$q \to p$} & &
      \colorbox{color2}{$p \to q$} & & \colorbox{color4}{$\neg p \lor \neg q$}
      \\
      \colorbox{color3}{$p$} \ar[urr,dash,gray] \ar[urrrr,dash,gray] & &
      \colorbox{color2}{$q$} \ar[u,dash,gray] \ar[urrrr,dash,gray] & &
      \colorbox{color2}{$p \leftrightarrow q$} \ar[u,dash,gray]
      \ar[urr,dash,gray] & & \colorbox{color4}{$p + q$} \ar[ullll,dash,gray]
      \ar[urr,dash,gray] & & \colorbox{color4}{$\neg q$} \ar[ullll,dash,gray]
      \ar[u,dash,gray] & & \colorbox{color1}{$\neg p$} \ar[ullll,dash,gray]
      \ar[ull,dash,gray] \\
      & & \fcolorbox{black}{color2}{$p \land q$} \ar[ull,dash,gray]
      \ar[u,dash,gray] \ar[urr,dash,gray] & & \fcolorbox{black}{color4}{$p \land
        \neg q$} \ar[ullll,dash,gray] \ar[urr,dash,gray] \ar[urrrr,dash,gray] &
      & \fcolorbox{black}{color1}{$\neg p \land q$} \ar[ullll,dash,gray]
      \ar[u,dash,gray] \ar[urrrr,dash,gray] & & \fcolorbox{black}{color1}{$\neg
        p \land \neg q$} \ar[ullll,dash,gray] \ar[u,dash,gray]
      \ar[urr,dash,gray] \\
      & & & & & \colorbox{color1}{$0$} \ar[ulll,dash,gray] \ar[ul,dash,gray]
      \ar[ur,dash,gray] \ar[urrr,dash,gray] & & &
    \end{tikzcd}
  \]
  \[
    \begin{tikzcd}
      & \colorbox{color3}{$\left[1\right]$} \ar[dl,dash,gray] \ar[dr,dash,gray]
      & \\
      \fcolorbox{black}{color2}{$\left[q\right]$} & &
      \fcolorbox{black}{color4}{$\left[\neg q\right]$} \\
      & \colorbox{color1}{$\left[0\right]$} \ar[ul,dash,gray] \ar[ur,dash,gray]
      &
    \end{tikzcd}
  \]
  \caption{Two BAs from \cref{example:construction}: $B(\mathcal{L})$ at the top
    and $B(\Delta)$ at the bottom. An edge between elements $a$ and $b$ (with
    $a$ positioned lower than $b$) means that $a < b$. Each element of
    $B(\Delta)$ is an equivalence class of elements of $B(\mathcal{L})$, and the
    colours show which elements of $B(\mathcal{L})$ belong to which class. In
    both algebras, atoms have borders around them.}
  \label{fig:example}
\end{figure}

@@

  The corresponding BA $B(\Delta)$ can then be constructed using
  \cref{def:algebra_from_logic}. Alternatively, one can first construct the free
  BA generated by the set $\{ p, q \}$---this corresponds to $B(\mathcal{L})$ in
  \cref{fig:example}---and then take a quotient with respect to either the
  filter generated by $p$ or the ideal\footnote{More details on these concepts
    can be found in many books on BAs
    \cite{givant2008introduction,koppelberg1989handbook}.} generated by $\neg
  p$. In any case, the resulting BA is pictured at the bottom of
  \cref{fig:example}.

  Each element of $B(\mathcal{L})$ can also be seen as a subset of the set of
  all models of $\mathcal{L}$, with $0$ representing $\emptyset$, $1$
  representing the set of all (four) models, each atom representing a single
  model, and each edge going upward representing a subset relation. Thus,
  the Boolean-algebraic way of calculating the WMC of $\Delta$ consists of:
  \begin{enumerate} % TODO: why is step 1 always possible?
  \item Identifying an element $a \in B(\mathcal{L})$ that corresponds to
    $\Delta$.
  \item Finding all atoms of $B(\mathcal{L})$ that are `dominated' by $a$
    according to the partial order.
  \item Using $w$ to calculate the weight of each such atom.
  \item Adding the weights of these atoms.
  \end{enumerate}
  This motivates the following definition of WMC generalised to BAs.
\end{example}
% TODO: how to compute the number of elements in the algebra.
% TODO: clarify what B(L) means. And whether B(Delta) is even necessary.
% TODO: reference for the set/subset thing.

% TODO: Feedback 2: You need to explain what precisely these mean in logic and
% models and weight functions are usually defined and understood.
% TODO: This should be replaced with inner sums (a.k.a. free products)
\begin{definition} \label{def:wmc}
  Let $\mathbf{B}$ be an atomic BA, and let $M \subset \mathbf{B}$ be its set of
  atoms. Let $L \subset \mathbf{B}$ be such that every atom $m \in M$ can be
  uniquely expressed as $m = \bigwedge L'$ for some $L' \subseteq L$, and let
  $w\colon L \to \mathbb{R}_{\ge 0}$ be arbitrary. The \emph{weighted model
    count} $\WMC_w\colon \mathbf{B} \to \mathbb{R}_{\ge 0}$ is defined as
  \[
    \WMC_w(x) = \begin{cases}
      0 & \text{if } x = 0 \\
      \prod_{l \in L'} w(l) & \text{if } M \ni x = \bigwedge L' \\
      \sum_{\text{atoms } a \le x} \WMC_w(a) & \text{otherwise}
    \end{cases}
  \]
  for any $x \in \mathbf{B}$. Furthermore, we define the \emph{normalised
    weighted model count} $\nWMC_w\colon \mathbf{B} \to [0, 1]$ as $\nWMC_w(x) =
  \frac{\WMC_w(x)}{\WMC_w(1)}$ for all $x \in \mathbf{B}$. For both $\WMC_w$ and
  $\nWMC_w$, we will drop the subscript when doing so results in no potential
  confusion.
%  Finally, we say that a measure $m\colon \mathbf{B} \to
%  \mathbb{R}_{\ge 0}$ is a \emph{WMC measure} (or is \emph{WMC-computable}) if
%  there exists a subset $L \subset \mathbf{B}$ and a weight function $w\colon L
%  \to \mathbb{R}_{\ge 0}$ such that $m = \WMC_w$.
\end{definition}
% TODO: mention that the definition can be reduced to a single formula (i.e.,
% without cases)
% TODO: any measure is a WMC measure if all atoms are in L

\begin{theorem}
  $\WMC$ is a measure, and $\nWMC$ is a probability measure.
\end{theorem}
% TODO: Feedback 2: Need to explain how WMC and NWMC connects to standard
% definitions of WMC and Pr(phi|e) = WMC(phi /\ e)/WMC(e).
\begin{proof}
  First, note that $\WMC$ is non-negative and $\WMC(0) = 0$ by definition. Next,
  let $x, y \in \mathbf{B}$ be such that $x \land y = 0$. We want to show that
  \begin{equation} \label{eq:additivity_proof}
    \WMC(x \lor y) = \WMC(x) + \WMC(y).
  \end{equation}
  If, say, $x = 0$, then \cref{eq:additivity_proof} becomes
  \[
    \WMC(y) = \WMC(0) + \WMC(y) = \WMC(y)
  \]
  (and likewise for $y = 0$). Thus we can assume that $x \ne 0 \ne y$ and use
  \cref{thm:representation} to write
  \[
    x = \bigvee_{i \in I} x_i \quad \text{and} \quad y = \bigvee_{j \in J} y_j
  \]
  for some sequences of atoms $(x_i)_{i \in I}$ and $(y_j)_{j \in J}$. If
  $x_{i'} = y_{j'}$ for some $i' \in I$ and $j' \in J$, then
  \[
    x \land y = \bigvee_{i \in I} \bigvee_{j \in J} x_i \land y_j = x_{i'} \land
    y_{j'} \ne 0,
  \]
  contradicting the assumption. This is enough to show that
  \begin{align*}
    \WMC(x \lor y) &= \WMC\left( \left( \bigvee_{i \in I} x_i \right) \lor \left(\bigvee_{j \in J} y_j \right) \right) = \sum_{i \in I} \WMC(x_i) + \sum_{j \in J} \WMC(y_j) \\
                   &= \WMC(x) + \WMC(y),
  \end{align*}
  finishing the proof that $\WMC$ is a measure. This immediately shows that
  $\nWMC$ is a probability measure since, by definition, $\nWMC(1) = 1$.
\end{proof}

@@

% TODO: Feedback 2: I [???] the significance of the below results. Proving
% properties that don't [???] the "roadmap" means (AI) readers will lose
% interest. [This applies until the beginning of the next section.]

\begin{lemma} \label{lemma:sum}
  For any measure $m\colon \mathbf{B} \to \mathbb{R}_{\ge 0}$ and elements $a, b
  \in \mathbf{B}$,
  \[
    m(a \lor b) = m(a) + m(b) - m(a \land b).
  \]
\end{lemma}
\begin{proof}
  By \cref{def:measure},
  \begin{align*}
    m(a) &= m(a \land b) + m(a \land \neg b), \\
    m(b) &= m(a \land b) + m(\neg a \land b), \\
    m(a \lor b) &= m(a \land b) + m(a \land \neg b) + m(\neg a \land b),
  \end{align*}
  so
  \[
    m(a) + m(b) - m(a \land b) = m(a \land b) + m(a \land \neg b) + m(\neg a
    \land b) = m(a \lor b)
  \]
  as required.
\end{proof}

\begin{lemma} \label{lemma:well-defined}
  For any $a, b \in \mathbf{B}$ and any principal ideal $(p)$, if $a/(p) =
  b/(p)$, then $a \lor p = b \lor p$.
\end{lemma}
\begin{proof}
  Note that
  \[
    a/(p) = b/(p) \quad \iff \quad a+b \in (p) \quad \iff \quad a+b \le p \quad
    \iff \quad (a+b) \lor p = p
  \]
  by \cref{def:quotient,def:ideal}, and the definition of $\le$. So $p = (a
  \land \neg b) \lor (\neg a \land b) \lor p$, and thus
  \[
    0 = p \land \neg p = (a \land \neg b \land \neg p) \lor (\neg a \land b
    \land \neg p) \lor (p \land \neg p) = (a \land \neg (b \lor p)) \lor (b
    \land \neg (a \lor p)).
  \]
  It follows that
  \[
    a \land \neg (b \lor p) = 0 \quad \text{and} \quad b \land \neg (a \lor p) =
    0.
  \]
  Focusing on the first equation,
  \[
    \neg a = (\neg a \lor a) \land [\neg a \lor \neg (b \lor p)] = \neg [a \land
    (b \lor p)],
  \]
  and so $a = a \land (b \lor p)$, and
  \[
    a \lor p = (a \lor p) \land (b \lor p) = (a \land b) \lor p.
  \]
  By similar arguments, $b \lor p = (a \land b) \lor p$ as well which shows that
  $a \lor p = b \lor p$ as required.
\end{proof}

% TODO: Feedback 2: Where is this helping?
\todo[inline]{Outdated. $m(a \land \neg p)$ is better than $m (a \lor p)$.}
\begin{proposition}[Measures on quotients] \label{def:measures_on_quotients}
  Let $m\colon \mathbf{B} \to \mathbb{R}_{\ge 0}$ be a measure, and let $(p)$ be
  a principal ideal. Let $m^*\colon \mathbf{B}/(p) \to \mathbb{R}_{\ge 0}$ be
  defined as
  \[
    m^*(a/(p)) = m(a \lor p)
  \]
  for any $a \in \mathbf{B}$. The function $m^*$ is well-defined. Furthermore,
  it is a measure on $\mathbf{B}/(p)$ if and only if $m(p) = 0$. Moreover, if it
  is a measure, then the following properties transfer from $m$ to $m^*$:
  \begin{itemize}
  \item if $m$ is a \emph{probability measure}, then so is $m^*$;
  \item if $m$ is \emph{strictly positive}, then so is $m^*$.
  \end{itemize}
\end{proposition}
\begin{proof}
  \Cref{lemma:well-defined} proves that the function is well-defined. Next,
  note that
  \[
    m^*(0/(p)) = m(0 \lor p) = m(p),
  \]
  so $m^*(0/(p)) = 0$ if and only if $m(p) = 0$. For the second part of
  \cref{def:measure}, let $a/(p), b/(p) \in \mathbf{B}/(p)$ be such that
  \[
    a/(p) \land b/(p) = a \land b / (p) = 0 / (p).
  \]
  This condition is equivalent to $a \land b \in (p)$ and $(a \land b) \lor p =
  p$ by well-known properties of quotients and ideals
  \cite{givant2008introduction}, \cref{def:ideal}, and the definition of $\le$,
  respectively. Now
  \begin{align*}
    m^*(a/(p) \lor b/(p)) &= m^*(a \lor b / (p)) = m(a \lor b \lor p) = m((a \lor p) \lor (b \lor p)) \\
                          &= m(a \lor p) + m(b \lor p) - m((a \lor p) \land (b \lor p)) \\
                          &= m^*(a/(p)) + m^*(b/(p)) - m((a \lor p) \land (b \lor p))
  \end{align*}
  by \cref{lemma:sum}. However
  \[
    (a \lor p) \land (b \lor p) = (a \land b) \lor p = p,
  \]
  so $m^*(a/(p) \lor b/(p)) = m^*(a/(p)) + m^*(b/(p))$ if and only if $m(p) = 0$.

  The two remaining properties are easy to prove:
  \begin{itemize}
  \item If $m(1) = 1$, then $m^*(1/(p)) = m(1 \lor p) = m (1) = 1$.
  \item Suppose that $m$ is strictly positive, and let $a/(p) \in
    \mathbf{B}/(p)$ be such that $a/(p) \ne 0/(p)$. Then
    \[
      m^*(a/(p)) = m(a \lor p) \ge m(a) > 0,
    \]
    where the first inequality comes from an elementary property of $\le$ that
    $x \le x \lor y$ for any $x, y \in \mathbf{B}$ \cite{sikorski1969boolean}
    and \cref{lemma:measure_and_order}; and the second inequality follows
    because $a/(p) \ne 0/(p)$ implies that $a \ne 0$, and $m$ is assumed to be
    strictly positive.
  \end{itemize}
\end{proof}
% TODO: I think I need to cite something for: a/I = 0/I <=> a \in I

\subsection{Lemma Galore}

\todo[inline]{This section made me realise that I was using the wrong definition}

% TODO: maybe 'unite' this with the result about a \/ p, i.e., a ~ a \/ p ~ a /\
% ~p.
% TODO: maybe I should state that A,B are assumed to be BAs with no additional
% properties unless stated otherwise.
\begin{lemma} \label{lemma:elements_in_quotients}
  Let $(p)$ be a principal ideal. Then for any $a \in \mathbf{B}$, $(a \land
  \neg p)/(p) = a/(p)$.
\end{lemma}
\begin{proof}
  Note that
  \[
    (a \land \neg p)/(p) = a/(p) \quad \iff \quad (a \land \neg p) + a \in (p)
    \quad \iff \quad (a \land \neg p) + a \le p.
  \]
  We also have that
  \[
    (a \land \neg p) + a = (a \land \neg p \land \neg a) \lor (\neg(a \land \neg
    p) \land a) = (\neg a \lor p) \land a = (\neg a \land a) \lor (p \land a) =
    p \land a.
  \]
  And, since $p \land a \le p$, we have that $(a \land \neg p) + a \le p$ as
  required.
\end{proof} % TODO: why? (the last bit) It's elementary
% TODO: remove all those 'let (p) be a principal ideal'.

\begin{lemma} \label{lemma:min_element_and_order}
  Let $(p)$ be a principal ideal. For any $a, b \in \mathbf{B}$, $a/(p) \le
  b/(p)$ if and only if $a \land \neg p \le b \land \neg p$.
\end{lemma}
\begin{proof}
  Let us begin with the `only of' direction. \Cref{lemma:elements_in_quotients}
  tells us that $(a \land \neg p)/(p) = a/(p)$. Combining this with
  \cref{lemma:canonical_homomorphism,lemma:homomorphisms_and_order} shows that
  \[
    a \land \neg p \le b \land \neg p \quad \implies \quad (a \land \neg p)/(p)
    \le (b \land \neg p)/(p) \quad \iff \quad a/(p) \le b/(p)
  \]
  as required.

  For the other direction, let $a, b \in \mathbf{B}$ be such that $a/(p) \le
  b/(p)$. Then, by \cref{lemma:order},
  \[
    [a/(p)] \land \neg [b/(p)] = (a \land \neg b)/(p) = 0/(p),
  \]
  i.e.,
  \[
    a \land \neg b \in (p) \quad \iff \quad a \land \neg b \le p \quad \iff
    \quad a \land \neg b \land \neg p = 0
  \]
  by \cref{def:ideal,lemma:order}. We need to show that $a \land \neg p \le b
  \land \neg p$. By \cref{lemma:order}, this is equivalent to $a \land \neg p
  \land \neg(b \land \neg p) = 0$. But
  \[
    a \land \neg p \land \neg(b \land \neg p) = a \land \neg p \land (\neg b
    \lor p) = (a \land \neg p \land \neg b) \lor (a \land \neg p \land p) = a
    \land \neg p \land \neg b,
  \]
  and we already have that $a \land \neg p \land \neg b = 0$ by assumption.
\end{proof}

\begin{lemma} \label{lemma:measure_of_atom}
  Let $m\colon \mathbf{B} \to \mathbb{R}_{\ge 0}$ be a measure, let $p \in
  \mathbf{B}$ be such that $m(p) = 0$, and let $m^*\colon \mathbf{B}/(p) \to
  \mathbb{R}_{\ge 0}$ be a measure defined by $m^*(a/(p)) = m(a \lor p)$. Then
  for any $a \in \mathbf{B}$, if $a/(p)$ is an atom in $\mathbf{B}/(p)$, then $a
  \land \neg p$ is an atom in $\mathbf{B}$ such that $m^*(a/(p)) = m(a \land
  \neg p)$.
\end{lemma}
\begin{proof} % TODO: This proof is a mess and needs more citations.
  First, we want to show that if $a/(p)$ is an atom, then $a \land \neg p$ is an
  atom. We can instead prove the contrapositive statement, i.e., if there exists
  a $b \in \mathbf{B}$ such that $0 < b < a \land \neg p$, then there exists a
  $b' \in \mathbf{B}$ such that $0/(p) < b'/(p) < a/(p)$. We will show that, in
  fact, we set $b' = b$.
  \Cref{lemma:canonical_homomorphism,lemma:homomorphisms_and_order} already tell
  us that $b/(p) \le a/(p)$, so we only need to show that $0/(p) < b/(p) \ne
  a/(p)$. For the first part, note that
  \[
    0/(p) < b/(p) \quad \iff \quad b/(p) \ne 0/(p) \quad \iff \quad b \not\in
    (p) \quad \iff \quad b \not\le p \quad \iff \quad b \land \neg p \ne 0
  \]
  by \cref{lemma:order}. But if $b \land \neg p = 0$, then $b \land a \land \neg
  p = 0$. This contradicts either that $b \le a \land \neg p$ (i.e., $b \land a
  \land \neg p = b$) or that $b \ne 0$. For the second part, i.e., $b/(p) \ne
  a/(p)$, we will show that if $b/(p) = a/(p)$, and $b \le a \land \neg p$, then
  $b = a \land \neg p$. Indeed,
  \[
    b/(p) = a/(p) \quad \iff \quad a + b \in (p) \quad \iff \quad a+b \le p
    \quad \iff \quad (a+b) \land \neg p = 0,
  \]
  and
  \[
    (a+b) \land \neg p = [(a \land \neg b) \lor (\neg a \land b)] \land \neg p =
    (a \land \neg b \land \neg p) \lor (\neg a \land b \land \neg p),
  \]
  so $(a+b) \land \neg p = 0$ implies that $a \land \neg b \land \neg p = 0$
  which is equivalent to $a \land \neg p \le b$. Therefore we have that $a \land
  \neg p \le b \le a \land \neg p$, so $b = a \land \neg p$ which, by
  contradiction, shows that $b/(p) \ne a/(p)$ and finishes the proof that $0/(p)
  < b/(p) < a/(p)$.

  In order to show that $m^*(a/(p)) = m(a \land \neg p)$, note that $a \land p$,
  $a \land \neg p$, and $\neg a \land p$ are pairwise disjoint and their
  supremum is $a \lor p$, so we have that
  \[
    m^*(a/(p)) = m(a \lor p) =  m(a \land p) + m(a \land \neg p) + m(\neg a
    \land p).
  \]
  Furthermore, since $a \land p \le p$, $m(a \land p) \le m(p) = 0$. Similarly,
  $m(\neg a \land p) = 0$, so $m^*(a/(p)) = m(a \land \neg p)$ as required.
\end{proof}
% TODO: the undergraduate book vaguely mentions that, when taking a quotient
% w.r.t. an ideal of elements with measure zero, all elements in an equivalence
% class will have the same measure.

>>>>

\newtheorem{theorem}{Theorem}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{conjecture}[theorem]{Conjecture}
\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{example}[theorem]{Example}
\theoremstyle{remark}
\newtheorem*{remark}{Remark}

@@

% Feedback: Can you say something here about factorized vs non-factorized
% weight function definitions? That is, factorized is when w maps literals to
% R_>=0, non-factorized is when w maps models to R_>=0 and show:
% a) come up with nice example when non-factorized weights are intuitive
% c) clarify that the factorized definition have is w.r.t. models, in case some
% one gets confused [It doesn't have to be, if the BA is not free -- P.]

@@

% TODO for later:
% 1) make up my mind about a,b vs x,y and stick to it (maybe x,y?)

@@

% TODO
% 1) Feedback 2: Explain with examples: models = elements [atoms] of algebra
% 2) Perhaps reorder the section of preliminaries into paragraphs, i.e., a
% paragraph for order, for homomorphisms, etc. This would take up less space.
% 3) how to compute the number of elements in the algebra.

@@

For clarity and succinctness, we will occasionally use three other operations
that can be defined using the original three\footnote{We use $+$ to denote
  symmetric difference because it is the additive operation of a Boolean ring.}:
\begin{align*}
  a \to b &= \neg a \lor b, \\
  a \leftrightarrow b &= (a \land b) \lor (\neg a \land \neg b), \\
  a + b &= (a \land \neg b) \lor (\neg a \land b).
\end{align*}
We can also define a partial order $\le$ on $\mathbf{B}$ as $a \le b$ if $a = b
\land a$ (or, equivalently, $a \lor b = b$) for all $a, b \in \mathbf{B}$.
Furthermore, let $a < b$ denote $a \le b$ and $a \ne b$. For the rest of this
paper, let $\mathbf{B}$ refer to the BA $(\mathbf{B}, \land, \lor, \neg, 0, 1)$.
For any $S \subseteq \mathbf{B}$, we write $\bigvee S$ for $\bigvee_{x \in S} x$
and call it the \emph{supremum} of $S$. Similarly, $\bigwedge S = \bigwedge_{x
  \in S} x$ is the \emph{infimum}. By convention, $\bigwedge \emptyset = 1$ and
$\bigvee \emptyset = 0$. For any $a, b \in \mathbf{B}$, we say that $a$ and $b$
are \emph{disjoint} if $a \land b = 0$.

@@

@@

@@

\section{WMC as a Measure}
% TODO: Feedback 2: Need to explain how WMC and NWMC connects to standard
% definitions of WMC and Pr(phi|e) = WMC(phi /\ e)/WMC(e).

@@

@@

  % TODO: be careful about mentioning ideals, filters, and quotients
  The corresponding BA $B(\Delta)$ can then be constructed using
  \cref{def:algebra_from_logic}. Alternatively, one can first construct the free
  BA generated by the set $\{ p, q \}$ and then take a quotient with respect to
  either the filter generated by $p$ or the ideal\footnote{More details on these
    concepts can be found in many books on BAs
    \cite{givant2008introduction,koppelberg1989handbook}.} generated by $\neg
  p$.

  Each element of $B(\mathcal{L})$ can also be seen as a subset of the set of
  all models of $\mathcal{L}$, with $0$ representing $\emptyset$, $1$
  representing the set of all (four) models, each atom representing a single
  model, and each edge going upward representing a subset relation. Thus,
  the Boolean-algebraic way of calculating the WMC of $\Delta$ consists of:
  \begin{enumerate} % TODO: why is step 1 always possible?
  \item Identifying an element $a \in B(\mathcal{L})$ that corresponds to
    $\Delta$.
  \item Finding all atoms of $B(\mathcal{L})$ that are `dominated' by $a$
    according to the partial order.
  \item Using $w$ to calculate the weight of each such atom.
  \item Adding the weights of these atoms.
  \end{enumerate}
  This motivates the following definition of WMC generalised to BAs.
\end{example}
% TODO: clarify what B(L) means. And whether B(Delta) is even necessary.
% TODO: reference for the set/subset thing.

% TODO: Feedback 2: You need to explain what precisely these mean in logic and
% models and weight functions are usually defined and understood.
% TODO: This should be replaced with inner sums (a.k.a. free products)
\begin{definition} \label{def:wmc}
  Let $\mathbf{B}$ be an atomic BA, and let $M \subset \mathbf{B}$ be its set of
  atoms. Let $L \subset \mathbf{B}$ be such that every atom $m \in M$ can be
  uniquely expressed as $m = \bigwedge L'$ for some $L' \subseteq L$, and let
  $w\colon L \to \mathbb{R}_{\ge 0}$ be arbitrary. The \emph{weighted model
    count} $\WMC_w\colon \mathbf{B} \to \mathbb{R}_{\ge 0}$ is defined as
  \[
    \WMC_w(x) = \begin{cases}
      0 & \text{if } x = 0 \\
      \prod_{l \in L'} w(l) & \text{if } M \ni x = \bigwedge L' \\
      \sum_{\text{atoms } a \le x} \WMC_w(a) & \text{otherwise}
    \end{cases}
  \]
  for any $x \in \mathbf{B}$. Furthermore, we define the \emph{normalised
    weighted model count} $\nWMC_w\colon \mathbf{B} \to [0, 1]$ as $\nWMC_w(x) =
  \frac{\WMC_w(x)}{\WMC_w(1)}$ for all $x \in \mathbf{B}$. For both $\WMC_w$ and
  $\nWMC_w$, we will drop the subscript when doing so results in no potential
  confusion. Finally, we say that a measure $m\colon \mathbf{B} \to
  \mathbb{R}_{\ge 0}$ is a \emph{WMC measure} (or is \emph{WMC-definable}) if
  there exists a subset $L \subset \mathbf{B}$ and a weight function $w\colon L
  \to \mathbb{R}_{\ge 0}$ such that $m = \WMC_w$.
\end{definition}
% TODO: mention that the definition can be reduced to a single formula (i.e.,
% without cases)
% TODO: any measure is a WMC measure if all atoms are in L

\begin{theorem}
  $\WMC$ is a measure, and $\nWMC$ is a probability measure.
\end{theorem}
\begin{proof}
  First, note that $\WMC$ is non-negative and $\WMC(0) = 0$ by definition. Next,
  let $x, y \in \mathbf{B}$ be such that $x \land y = 0$. We want to show that
  \begin{equation} \label{eq:additivity_proof}
    \WMC(x \lor y) = \WMC(x) + \WMC(y).
  \end{equation}
  If, say, $x = 0$, then \cref{eq:additivity_proof} becomes
  \[
    \WMC(y) = \WMC(0) + \WMC(y) = \WMC(y)
  \]
  (and likewise for $y = 0$). Thus we can assume that $x \ne 0 \ne y$ and use
  \cref{thm:representation} to write
  \[
    x = \bigvee_{i \in I} x_i \quad \text{and} \quad y = \bigvee_{j \in J} y_j
  \]
  for some sequences of atoms $(x_i)_{i \in I}$ and $(y_j)_{j \in J}$. If
  $x_{i'} = y_{j'}$ for some $i' \in I$ and $j' \in J$, then
  \[
    x \land y = \bigvee_{i \in I} \bigvee_{j \in J} x_i \land y_j = x_{i'} \land
    y_{j'} \ne 0,
  \]
  contradicting the assumption. This is enough to show that
  \begin{align*}
    \WMC(x \lor y) &= \WMC\left( \left( \bigvee_{i \in I} x_i \right) \lor \left(\bigvee_{j \in J} y_j \right) \right) = \sum_{i \in I} \WMC(x_i) + \sum_{j \in J} \WMC(y_j) \\
                   &= \WMC(x) + \WMC(y),
  \end{align*}
  finishing the proof that $\WMC$ is a measure. This immediately shows that
  $\nWMC$ is a probability measure since, by definition, $\nWMC(1) = 1$.
\end{proof}

@@

\section{What Measures Are WMC-Definable?}
